Relative Stable Maps in Genus One via Central Alignments

### 2018-02-10 genus-one-relative-maps.tex
Finished example of expanded degeneration justifying novel condition (it ended up quite long)

\newcommand{\TT}{\operatorname{T}}
\newcommand{\oM}{\overline{\mathcal{M}}}
\newcommand{\M}[4]{\overline{\mathcal{M}}_{#1,#2}(#3,#4)}
\newcommand{\Q}[4]{\mathcal{Q}_{#1,#2}(#3,#4)}
\newcommand{\Qe}[4]{\mathcal{Q}^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\Qt}[4]{\widetilde{\mathcal Q}_{#1,#2}(#3,#4)}
\newcommand{\QG}[4]{\mathcal{Q}G_{#1,#2}(#3,#4)}
\newcommand{\QGe}[4]{\mathcal{Q}G^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\D}[3]{\mathcal{D^Q}(#1,#2,#3)}
\newcommand{\E}[3]{\mathcal{E^Q}(#1,#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\tVZc}[4]{\widetilde{\mathcal{V\!Z}}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZc}[4]{\mathcal{V\!Z}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZcLi}[4]{\mathcal{V\!Z}^{\rm{ctr, Li}}_{#1,#2}(#3,#4)}
\newcommand{\N}{\mathbb{N}}
\newcommand{\OO}{\mathcal{O}}
\renewcommand{\to}{\rightarrow}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\EE}{\mathbf{E}}
\renewcommand{\L}{\mathcal L}
\newcommand{\LL}{\mathbf{L}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\Aaff}{\mathbb{A}}
\newcommand{\kfield}{\Bbbk}
\newcommand{\comp}{\chi}
\newcommand{\sst}{\sigma^{\operatorname{ss}}}
\newcommand{\Pic}{\operatorname{Pic}}
\newcommand{\Def}{\operatorname{Def}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Proj}{\operatorname{Proj}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gm}{\mathbb{G}_{\text{m}}}
\newcommand{\virt}[1]{[#1]^{\operatorname{virt}}}
\newcommand{\vip}[1]{[#1]^{\operatorname{prod}}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\operatorname{H}}
\newcommand{\Achow}{\operatorname{A}}
\newcommand{\pt}{\operatorname{pt}}
\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\ba}{\begin{aligned}}
\newcommand{\ea}{\end{aligned}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bsm}{\left(\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}\right)}                   
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}
\newcommand{\barr}{\begin{displaymath}\begin{array}{cccc}}
\newcommand{\earr}{\end{array}\end{displaymath}}
\newcommand{\barrl}{\begin{displaymath}\begin{array}{lcl}}
\newcommand{\earrl}{\end{array}\end{displaymath}}
\newcommand{\barl}{\begin{displaymath}\begin{array}{l}}
\newcommand{\earl}{\end{array}\end{displaymath}}
\newcommand{\bxym}{ \begin{displaymath}\xymatrix }
\newcommand{\exym}{\end{displaymath}}
\newcommand{\bcd}{\begin{center}\begin{tikzcd}}
\newcommand{\ecd}{\end{tikzcd}\end{center}}
\newcommand{\R}{\operatorname{R}^{\bullet}}
%\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\Isom}{\text{Isom}}
\newcommand{\pr}{\operatorname{pr}}
\newcommand{\ev}{\operatorname{ev}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\vdim}{\operatorname{vdim}}
\newcommand{\ildef}[1]{\emph{#1}}
\newcommand{\om}[1]{\mathcal{#1}}
\newcommand{\h}{\operatorname{h}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\RR}{\textbf{R}}
\newcommand{\NN}{\operatorname{N}}

@@

Consider now the following weighted graph for a map in the boundary
\begin{center}
\begin{tikzpicture}
\draw[color=brown] (0,0) node[left]{$E$} -- (4,0);
\draw[fill=black] (1,0) circle[radius=2pt] node[above]{$x_1$};
\draw[fill=black] (2,0) circle[radius=2pt] node[below right]{$y_1$};
\draw[fill=black] (3,0) circle[radius=2pt] node[below right]{$y_2$};
\draw (2,1.5) node[left]{$R_1$} (3,1.5) node[left]{$R_2$};
\draw (2,-1) -- (2,2) node[above]{$2$} (3,-1) -- (3,2) node[above]{$1$};
\draw[->] (5,0.5) -- (8,0.5);
\draw (9,-1) -- (9,2) node[above]{$\PP^1$};
\draw[fill=black] (9,0) circle[radius=2pt] node[right]{$H$};
\end{tikzpicture}
\end{center}
where the brown line represents a contracted genus $1$ curve. Now $(E,x_1,y_1,y_2)$ is a point of $\oM_{1,3}$ subject to the divisorial condition $3x_1-2y_1-y_2=0\in \Achow_0(E)$; furthermore we have to choose the second branch point of the $2\colon 1$ map from $R_1$ to $\PP^1$. This already makes up for a $3$-dimensional moduli space of degenerate relative maps corresponding to such a graph. The minimal log structure for this curve has a chart from $\mathbb N^2$, with generators $e_1$ and $e_2$ corresponding to the smoothing parameters of the two nodes. If we allowed $e_1$ and $e_2$ to be identified in the characteristic sheaf, then we would get an extra $\Gm$ of choices for the log structure, so in total a $4$-dimensional moduli space. Thus if we don't impose the novel condition then we get a whole other component of the relative space, which of course for dimensional reasons cannot be contained in the closure of the nice locus.
\end{example}

@@

So, let us pretend that we have a good definition of the ``centrally aligned Li space'' of maps from centrally aligned curves to expanded degenerations:
\begin{equation*} \VZcLi{1}{\alpha}{X|D}{\beta} \end{equation*}
Objects of this moduli space should consist of maps $C \to \mathfrak{X}$ where $\mathfrak{X} \to X$ is an expansion of the pair $(X,D)$, together with a central alignment on $C$ which is compatible with this map. The factorisation property should say that the following diagram commutes:
\bcd
& \tilde{C} \ar[ld,"\nu" left] \ar[rd,"\tau"] & \\
C \ar[d,"f" left] & & \overline{C} \ar[d,"\overline{f}", dotted] \\
\mathfrak{X} \ar[rd,"\pi"] & & \mathfrak{X} \ar[ld,"\pi"] \\
& X & 
\ecd
Morally speaking, the projection map
\begin{equation*} \pi_* \colon \VZcLi{1}{\alpha}{X|D}{\beta} \to \VZc{1}{n}{X}{\beta} \end{equation*}
should have as its image our space $\VZc{1}{\alpha}{X|D}{\beta}$. We will see that everything in the image of this space must satisfy the novel condition, thus providing further justification for us imposing it.

@@

First we have to identify what the possible lifts of this object to the centrally aligned Li space are. The map $C \to \mathfrak{X}$ takes the following form:

@@

\draw[red,very thick] (-0.5,-1) to [out=90,in=180] (-0.25,-0.8) to [out=0,in=90] (0,-1);
\draw[red,very thick] (-0.25,-0.8) -- (0,1);
\draw[red] (0.1,0) node{$E$};
\draw[red,fill=red] (0,1) circle[radius=2pt] node[above]{$5$};
\end{tikzpicture}
\end{center}
(Here the box at level $1$ represents a single fibre of the expanded degeneration; $E$ is mapped into a fibre and has zero horizontal degree.) In this case $\tilde{C}$ is obtained from $C$ by bubbling a rational component at the marking, and so the map $\tilde{C} \to \mathfrak{X}$ is given by:

\begin{center}
\begin{tikzpicture}
\draw (-1,-1) -- (1,-1);
\draw (-1,-1) -- (-1,1);
\draw (1,-1) -- (1,1);
\draw (-1,1) -- (1,1);
\draw (-1,-1) -- (-2,-2.5);
\draw (-2,-2.5) -- (0,-2.5);
\draw (0,-2.5) -- (1,-1);
\draw (-1,1) -- (-1,3);
\draw (-1,3) -- (1,3);
\draw (1,3) -- (1,1);
\draw[red,very thick] (-2,-2.5) -- (-0.5,-1);
\draw[red] (-1,-1.5) node[right]{$C_1$};
\draw[red,fill=red] (-0.5,-1) circle[radius=2pt];
\draw[red] (-0.6,-1) node[above]{$2$};

@@

\draw[red, very thick] (0,1) -- (0,3);
\draw[red] (0,2) node[right]{$C_3$};
\draw[red,fill=red] (0,3) circle[radius=2pt] node[above]{$5$};
\end{tikzpicture}
\end{center}

\end{example}

\subsection{Proof}
We now want to show that the relative space in the centrally aligned setting is equal to the closure of the nice locus; irreducibility then follows immediately. We start with one direction.

\begin{lem}
The closure of the nice locus is contained in the relative space.
\end{lem}
\begin{proof}
We address the relative conditions one at a time. Notice that they are all obviously satisfied on the nice locus.

The factorisation property is closed: see \cite[Theorem 4.3]{RSPW}.

Gathmann's relative condition is closed: see \cite[Proposition 4.9]{Vre}.

>>>>

\newcommand{\TT}{\operatorname{T}}
\newcommand{\oM}{\overline{\mathcal{M}}}
\newcommand{\M}[4]{\overline{\mathcal{M}}_{#1,#2}(#3,#4)}
\newcommand{\Q}[4]{\mathcal{Q}_{#1,#2}(#3,#4)}
\newcommand{\Qe}[4]{\mathcal{Q}^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\Qt}[4]{\widetilde{\mathcal Q}_{#1,#2}(#3,#4)}
\newcommand{\QG}[4]{\mathcal{Q}G_{#1,#2}(#3,#4)}
\newcommand{\QGe}[4]{\mathcal{Q}G^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\D}[3]{\mathcal{D^Q}(#1,#2,#3)}
\newcommand{\E}[3]{\mathcal{E^Q}(#1,#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\tVZc}[4]{\widetilde{\mathcal{V\!Z}}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZc}[4]{\mathcal{V\!Z}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZcLi}[4]{\mathcal{V\!Z}^{\rm{ctr, Li}}_{#1,#2}(#3,#4)}
\newcommand{\N}{\mathbb{N}}
\newcommand{\OO}{\mathcal{O}}
\renewcommand{\to}{\rightarrow}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\EE}{\mathbf{E}}
\renewcommand{\L}{\mathcal L}
\newcommand{\LL}{\mathbf{L}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\Aaff}{\mathbb{A}}
\newcommand{\kfield}{\Bbbk}
\newcommand{\comp}{\chi}
\newcommand{\sst}{\sigma^{\operatorname{ss}}}
\newcommand{\Pic}{\operatorname{Pic}}
\newcommand{\Def}{\operatorname{Def}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Proj}{\operatorname{Proj}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gm}{\mathbb{G}_{\text{m}}}
\newcommand{\virt}[1]{[#1]^{\operatorname{virt}}}
\newcommand{\vip}[1]{[#1]^{\operatorname{prod}}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\operatorname{H}}
\newcommand{\Achow}{\operatorname{A}}
\newcommand{\pt}{\operatorname{pt}}
\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\ba}{\begin{aligned}}
\newcommand{\ea}{\end{aligned}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bsm}{\left(\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}\right)}                   
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}
\newcommand{\barr}{\begin{displaymath}\begin{array}{cccc}}
\newcommand{\earr}{\end{array}\end{displaymath}}
\newcommand{\barrl}{\begin{displaymath}\begin{array}{lcl}}
\newcommand{\earrl}{\end{array}\end{displaymath}}
\newcommand{\barl}{\begin{displaymath}\begin{array}{l}}
\newcommand{\earl}{\end{array}\end{displaymath}}
\newcommand{\bxym}{ \begin{displaymath}\xymatrix }
\newcommand{\exym}{\end{displaymath}}
\newcommand{\bcd}{\begin{center}\begin{tikzcd}}
\newcommand{\ecd}{\end{tikzcd}\end{center}}
\newcommand{\R}{\operatorname{R}^{\bullet}}
%\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\Isom}{\text{Isom}}
\newcommand{\pr}{\operatorname{pr}}
\newcommand{\ev}{\operatorname{ev}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\vdim}{\operatorname{vdim}}
\newcommand{\ildef}[1]{\emph{#1}}
\newcommand{\om}[1]{\mathcal{#1}}
\newcommand{\h}{\operatorname{h}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\RR}{\textbf{R}}
\newcommand{\NN}{\operatorname{N}}
\newcommand{\ovm}[1]{\overline{\mathcal{#1}}}
\newcommand{\ovt}[1]{\widetilde{\mathcal{#1}}}
\newcommand{\ov}[1]{\overline{#1}}

@@

Consider now the following weighted graph for a map in the boundary
\begin{center}
\begin{tikzpicture}
\draw[color=brown] (0,0) node[left]{$E$} -- (4,0);
\draw[fill=black] (1,0) circle[radius=2pt] node[above]{$x_1$};
\draw[fill=black] (2,0) circle[radius=2pt] node[below right]{$y_1$};
\draw[fill=black] (3,0) circle[radius=2pt] node[below right]{$y_2$};
\draw (2,1.5) node[left]{$R_1$} (3,1.5) node[left]{$R_2$};
\draw (2,-1) -- (2,2) node[above]{$2$} (3,-1) -- (3,2) node[above]{$1$};
\draw[->] (5,0.5) -- (8,0.5);
\draw (9,-1) -- (9,2) node[above]{$\PP^1$};
\draw[fill=black] (9,0) circle[radius=2pt] node[right]{$H$};
\end{tikzpicture}
\end{center}
where the brown line represents a contracted genus $1$ curve. Now $(E,x_1,y_1,y_2)$ is a point of $\oM_{1,3}$ subject to the divisorial condition $3x_1-2y_1-y_2=0\in \Achow_0(E)$; furthermore we have to choose the second branch point of the $2\colon 1$ map from $R_1$ to $\PP^1$. This already makes up for a $3$-dimensional moduli space of degenerate relative maps corresponding to such a graph. The minimal log structure for this curve has a chart from $\mathbb N^2$, with generators $e_1$ and $e_2$ corresponding to the smoothing parameters of the two nodes. If we allowed $e_1$ and $e_2$ to be identified in the characteristic sheaf, then we would get an extra $\Gm$ of choices for the log structure, so in total a $4$-dimensional moduli space. Thus if we don't impose the novel condition then we get a whole other component of the relative space, which of course for dimensional reasons cannot be contained in the closure of the nice locus.\qed
\end{example}

@@

So, let us pretend that we have a good definition of the ``centrally aligned Li space'' of maps from centrally aligned curves to expanded degenerations:
\begin{equation*} \VZcLi{1}{\alpha}{X|D}{\beta} \end{equation*}
Objects of this moduli space should consist of maps $C \to X_l$ where $X_l \to X$ is an expansion (of some length $l$) of the pair $(X,D)$, together with a central alignment on $C$ which is compatible with this map. The factorisation property should say that the following diagram commutes:
\bcd
& \tilde{C} \ar[ld,"\nu" left] \ar[rd,"\tau"] & \\
C \ar[d,"f" left] & & \overline{C} \ar[d,"\overline{f}", dotted] \\
X_l \ar[rd,"\pi"] & & X_l \ar[ld,"\pi"] \\
& X & 
\ecd
Morally speaking, the projection map
\begin{equation*} \pi_* \colon \VZcLi{1}{\alpha}{X|D}{\beta} \to \VZc{1}{n}{X}{\beta} \end{equation*}
should have as its image our space $\VZc{1}{\alpha}{X|D}{\beta}$. We will see that everything in the image of this space must satisfy the novel condition, thus providing further justification for us imposing it.

@@

First we have to identify what the possible lifts of this object to the centrally aligned Li space are. The map $C \to X_1$ takes the following form:

@@

\draw[red,very thick] (-0.5,-1) to [out=90,in=180] (-0.25,-0.8) to [out=0,in=90] (0,-1);
\draw[red,very thick] (-0.25,-0.8) -- (0,1);
\draw[red] (0.1,0) node{$E$};
\draw[red,fill=red] (0,1) circle[radius=2pt] node[above]{$5$};
\end{tikzpicture}
\end{center}
(Here we use Navid's convention for drawing maps to expanded degenerations, so that the box at level $1$ represents a \emph{single fibre} of the expanded degeneration, i.e. $E$ is mapped into a fibre and has zero horizontal degree.) Now, $\widetilde{C}$ is obtained from $C$ by bubbling a rational component $C_3$ at the single marking, and so the map $\widetilde{C} \to X_2$ is given by:
\begin{center}
\begin{tikzpicture}
\draw (-1,-1) -- (1,-1);
\draw (-1,-1) -- (-1,1);
\draw (1,-1) -- (1,1);
\draw (-1,1) -- (1,1);
\draw (-1,-1) -- (-2,-2.5);
\draw (-2,-2.5) -- (0,-2.5);
\draw (0,-2.5) -- (1,-1);
\draw (-1,1) -- (-1,3);
\draw (-1,3) -- (1,3);
\draw (1,3) -- (1,1);
\draw[red,very thick] (-2,-2.5) -- (-0.5,-1);
\draw[red] (-1,-1.5) node[right]{$C_1$};
\draw[red,fill=red] (-0.5,-1) circle[radius=2pt];
\draw[red] (-0.6,-1) node[above]{$2$};

@@

\draw[red, very thick] (0,1) -- (0,3);
\draw[red] (0,2) node[right]{$C_3$};
\draw[red,fill=red] (0,3) circle[radius=2pt] node[above]{$5$};
\end{tikzpicture}
\end{center}
The map $\widetilde{C} \to \overline{C}$ contracts the elliptic component $E$, and so the map $\overline{C} \to X_1$ \emph{should} take the form:
\begin{center}
\begin{tikzpicture}
\draw (-1,-1) -- (1,-1);
\draw (-1,-1) -- (-1,1);
\draw (1,-1) -- (1,1);
\draw (-1,1) -- (1,1);
\draw (-1,-1) -- (-2,-2.5);
\draw (-2,-2.5) -- (0,-2.5);
\draw (0,-2.5) -- (1,-1);
\draw[red,very thick] (-2,-2.5) -- (0,-1);
\draw[red] (-0.8,-1.7) node[below]{$\overline{C}_1$};
\draw[red,fill=red] (0,-1) circle[radius=2pt];

\draw[red,very thick] (0,-2.5) -- (0,-1);
\draw[red] (-0.05,-1.5) node[right]{$\overline{C}_2$};
\draw[red,fill=red] (0,-1) circle[radius=2pt];

\draw[red,very thick] (0,-1) -- (0,1);
\draw[red] (0.35,0) node{$\overline{C}_3$};
\draw[red,fill=red] (0,1) circle[radius=2pt] node[above]{$5$};
\end{tikzpicture}
\end{center}
We will now see that, in this example, the map drawn above cannot actually exist, and that this occurs precisely because we have violated the novel condition.

We start by choosing a one-parameter smoothing $\mathcal{C} \to \Aaff^1$ of the centrally aligned curve $C$. This induces one-parameter smoothings $\widetilde{\mathcal{C}}$ of $\widetilde{C}$ and $\overline{\mathcal{C}}$ of $\overline{C}$. Since the factorisation condition must also hold in families, we must have a map $\overline{\mathcal{C}} \to \mathfrak{X}$, whose general fibre maps to the space $X_0=X$ (with no expansion) and whose general fibre maps to $X_1=X \sqcup_D Y$.

We thus have a map $\ovm{C} \to \mathfrak{X}$, where $\mathfrak{X}$ is the total space of the expanded degeneration. This is obtained as the toric blow-up of $\PP^1 \times \Aaff^1$ at the point $(\infty,0)$. Thus it is also a smooth toric variety, with fan given by:
\begin{center}
\begin{tikzpicture}
\draw[fill=black] (0,0) circle[radius=2pt];
\draw[thick] (0,0) -- (2,0) node[above]{$\rho_1$};
\draw[thick] (0,0) -- (1.75,1.75) node[right]{$\rho_2$};
\draw[thick] (0,0) -- (0,2) node[left]{$\rho_3$};
\draw[thick] (0,0) -- (0,-1.75) node[left]{$\rho_4$};
\end{tikzpicture}
\end{center}
There is a map $\mathfrak{X} \to \Aaff^1$ whose general fibre is isomorphic to $X$ and whose central fibre is isomorphic to $X_1$. We see from the fan above that $\Pic(\mathfrak{X})$ is generated by $\OO_{\mathfrak{X}}(D_{\rho_i})$ for $i=1,\ldots,4$, subject to the relations:
\begin{align*}
\OO_{\mathfrak{X}}(D_{\rho_1}) \otimes \OO_{\mathfrak{X}}(D_{\rho_2}) & \cong \OO_{\mathfrak{X}} \\
\OO_{\mathfrak{X}}(D_{\rho_2}) \otimes \OO_{\mathfrak{X}}(D_{\rho_3}) & \cong \OO_{\mathfrak{X}}(D_{\rho_4})
\end{align*}
Choosing as $\{\OO_{\mathfrak{X}}(D_{\rho_1}), \OO_{\mathfrak{X}}(D_{\rho_3})\}$ our minimal set of generators, we see that a map $\overline{\mathcal{C}} \to \mathfrak{X}$ is given by the data of line bundes $L_1$ and $L_3$ on $\overline{\mathcal{C}}$ together with sections:
\begin{align*}
s_1 & \in \HH^0(\overline{\mathcal{C}},L_1) \\
s_2 & \in \HH^0(\overline{\mathcal{C}},L_1^{-1}) \\
s_3 & \in \HH^0(\overline{\mathcal{C}},L_3) \\
s_4 & \in \HH^0(\overline{\mathcal{C}},L_1^{-1}\otimes L_3)
\end{align*}
The divisors $D_{\rho_1}$ and $D_{\rho_2}$ are, respectively, the level-$0$ and level-$1$ pieces of the central fibre of $\mathfrak{X}$. Therefore we must have
\begin{align*} & L_1 \cong \OO_{\overline{\mathcal{C}}}(a_1 \ov{C}_1 + a_2 \ov{C}_2) \\
& L_1^{-1} \cong \OO_{\ov{C}}(a_3 \ov{C}_3)
\end{align*}
for some positive integers $a_1,a_2$ and $a_3$. Here we are using the fact that each $\ov{C}_i$ is a $\QQ$-Cartier divisor on $\ovm{C}$, so that (after possibly multipltying by some positive integer) the corresponding line bundle makes sense. Furthermore, the intersection multiplicites of $\ov{C}_1$ and $\ov{C}_2$ with the level-$1$ piece mean that we must have:
\begin{align*} \deg(L_1^{-1}|_{\ov{C}_1}) & = 2\\
\deg(L_1^{-1}|_{\ov{C}_2}) & = 3 \end{align*}
We thus obtain:
\begin{align*} a_3 (\ov{C}_1 \cdot \ov{C}_3)&  = 2 \\
a_3 (\ov{C}_2 \cdot \ov{C}_3)& = 3 \end{align*}
However, we claim that in fact $\ov{C}_1 \cdot \ov{C}_3 = \ov{C}_2 \cdot \ov{C}_3$, so that the above equations cannot hold. To calculate these intersection numbers, we pass to the semistable model of $\ovm{C}$. This is obtained by performing further blow-ups to $\ovt{C}$, resulting in a semistable model $\ovt{C}^{\operatorname{ss}}$of the following form:
\begin{center}
\begin{tikzpicture}
\draw (-3,-3) -- (-3,2);
\draw (-3,-2.5) node[left]{$E$};

\draw (-3.5,1.5) -- (-0.5,1.5);
\draw (-1.3,1.725) -- (0.5,0.5);
\draw[fill=black] (0.8,0.5) circle[radius=0.5pt];
\draw[fill=black] (0.9,0.5) circle[radius=0.5pt];
\draw[fill=black] (1,0.5) circle[radius=0.5pt];
\draw (3,1.725) -- (1.2,0.5);
\draw (1.9,0.9) node[right]{$R_1^1$};
\draw (2.5,1.5) -- (5,1.5);
\draw (4,1.5) node[below]{$R_1$};

\draw (-3.5,0) -- (-0.5,0);
\draw (-1.3,0.225) -- (0.5,-1);
\draw[fill=black] (0.8,-1) circle[radius=0.5pt];
\draw[fill=black] (0.9,-1) circle[radius=0.5pt];
\draw[fill=black] (1,-1) circle[radius=0.5pt];
\draw (3,0.225) -- (1.2,-1);
\draw (1.9,-0.6) node[right]{$R_2^1$};
\draw (2.5,0) -- (5,0);
\draw (4,0) node[below]{$R_2$};

\draw (-3.5,-1.5) -- (-0.5,-1.5);
\draw (-1.3,-1.275) -- (0.5,-2.5);
\draw[fill=black] (0.8,-2.5) circle[radius=0.5pt];
\draw[fill=black] (0.9,-2.5) circle[radius=0.5pt];
\draw[fill=black] (1,-2.5) circle[radius=0.5pt];
\draw (3,-1.275) -- (1.2,-2.5);
\draw (2.5,-1.5) -- (5,-1.5);
\draw (4,-1.5) node[below]{$R_3$};
\end{tikzpicture}
\end{center}
Here the component labeled $E$ is smooth elliptic and all the other components are smooth rational. The projection $\pi: \ovt{C}^{\operatorname{ss}} \to \ovm{C}$ sends each $R_i$ to $\ov{C}_i$ and contracts all the other components. The semistable model has the advantage that the total space is regular, so the intersections are precisely those that you ``see'' in the picture. In particular, any two adjacent components have intersection $1$, and the self-intersection of any component is equal to minus the number of adjacent components.

Now, $\ov{C}_i \cdot \ov{C}_j = (\pi^* \ov{C}_i)\cdot(\pi^* \ov{C}_j)$, so we want to show that:
\begin{equation*} (\pi^* \ov{C}_1)\cdot(\pi^*\ov{C}_3) = (\pi^* \ov{C}_2)\cdot(\pi^*\ov{C}_3) \end{equation*}
For each $i$, $\pi^*\ov{C}_i$ is equal to $R_i$ plus a non-negative linear combination of the components in $\ovt{C}^{\operatorname{ss}}$ contracted by $\pi$. By the projection formula, any component contracted by $\pi$ will multiply to zero with a class pulled back along $\pi$. Thus
\begin{equation*} (\pi^* \ov{C}_1)\cdot(\pi^*\ov{C}_3) = R_1 \cdot (\pi^*\ov{C}_3) = a_3(R_1^1) \end{equation*}
the coefficient of $R_1^1$ in $\pi^* \ov{C}_3$. Similarly
\begin{equation*} (\pi^* \ov{C}_2) \cdot (\pi^* \ov{C}_3) = a_3(R_2^1) \end{equation*}
the coefficient of $R_2^1$ in $\pi^* \ov{C}_3$. So we need to show that:
\begin{equation*} a_3(R_1^1) = a_3(R_1^2) \end{equation*}
Now, we have
\begin{equation*} 0 = R_1^1 \cdot (\pi^* \ov{C}_3) = a_3(R_1^2) - 2a_3(R_1^1) \end{equation*}
because $(R_1^1)^2=-2$. So $a_3(R_1^1) = a_3(R_1^2)/2$. Similarly:
\begin{equation*} 0 = R_1^2 \cdot (\pi^*\ov{C}_3) = a_3(R_1^1) - 2a_3(R_1^2) + a_3(R_1^3) = -3a_3(R_1^1) + a_3(R_1^3) \end{equation*}
So $a_3(R_1^1) = a_3(R_1^3)/3$. Continuing in this way, we find that
\begin{equation*} a_3(R_1^1) = a_3(E)/l_1 \end{equation*}
where $l_1$ is the length of the rational tail connecting $E$ to $R_1$. Similarly $a_3(R_1^2) = a_3(E)/l_2$. But by Smyth's balancing condition for the semistable model \cite[Proposition 2.11]{SMY1}, we must have $l_1=l_2$. This concludes the proof that $\ov{C}_1 \cdot \ov{C}_3 = \ov{C}_2 \cdot \ov{C}_3$, which shows the necessity of the novel condition in this example.

Notice that here we used the fact that the rational tails $\ov{C}_1$ and $\ov{C}_2$ are adjacent to the same irreducible component (namely $E$) of the core.\qed
\end{example}

\subsection{Proof that the relative space equals the closure of the nice locus.}
We now want to show that the relative space in the centrally aligned setting is equal to the closure of the nice locus; irreducibility then follows immediately. We start with one direction:

\begin{lem}
The closure of the nice locus is contained in the relative space.
\end{lem}
\begin{proof}
We address the relative conditions one at a time. Notice that they are all obviously satisfied on the nice locus. Also, Gathmann's relative condition is obviously closed; see \cite[Proposition 4.9]{Vre}.
